## Microbes to watch

If you're brand new to the microbiome, the variety of microbes can be pretty daunting. Let's look through specific microbes you're likely to see and compare them to the ones that were in my samples. 

In this first step, we'll just consider the microbes that are commonly found, without considering how to change them. Later, in the chapter on [experiments](#experiments), we'll show more about how you can _manipulate_ them.


### The most common phyla

Most westerners will find their sample dominated by two phyla: _Firmicutes_ and _Bacteroidetes_. These are so common in healthy people that it’s tempting to assume this is "natural" or "normal", but like much else with the microbiome, the situation is different outside the western world, a clue that it’s difficult to summarize a single individual’s microbiome as “good” or “bad.” It all depends.

Of the samples I've collected, here's the general idea of how much of each phylum:

```{r}
Firmicutes.people <- subset_taxa(subset_samples(people.phylum, Site == "gut"), Phylum %in% c("Firmicutes","Bacteroidetes"))

mhg_density_for_taxa2 <- function(ps,tax_name, username=NULL) {
  df <- data.frame(abundance=as.numeric(otu_table(ps)[tax_name])/10000,
                   user = sample_data(ps)$Username,
                   date = sample_data(ps)$Date)
  ggplot(data=df,
         aes(x=abundance)) + 
    stat_density() +
    ggtitle(paste("Abundance of",tax_name,"across hundreds of samples")) 
    
}

mhg_density_for_taxa(subset_samples(Firmicutes.people,Condition == "Healthy"),"Firmicutes")
mhg_density_for_taxa(subset_samples(Firmicutes.people,Condition == "Healthy"),"Bacteroidetes")
```


*Firmicutes* is usually the most abundant in uBiome tests of modern, urban people, and for good reason: it’s very good at sucking energy out of the types of high-fat, highly-processed food we eat in industrialized societies. For that reason, some scientists have claimed it might associate with obesity, since its presence might indicate a gut that is highly-optimized to pull as many calories as possible from food. On the other hand, it’s not hard to find skinny people with lots of Firmicutes, so this theory doesn’t always hold. I've seen it lower in vegetarians, and I suspect that maybe it reflects diet more than weight.


*Bacteroidetes* is not as good at digesting fat, so it sticks to high-fiber foods like those in vegetables, which is why vegetarians tend to have high levels of this phylum. It’s often most prominent outside the industrialized world, too, where people eat less meat. Often, though not always, when I see a sample with high _Bacteroidetes_ (i.e. higher than _Firmicutes_) it'll turn out the person is a vegetarian. Incidentally, don't confuse this phylum with the similar-sounding genus _Bacteroides_ [see below].

*Prevotella* is associated with carbohydrates, and you might think they're more common in people who eat lots of 

*Proteobacteria* If you’re an otherwise normal omnivore and see double-digit amounts of this it can be a sign that your body was recently under attack. That's because the Proteobacteria includes many pathogens like *E. Coli* and *Shigella*. For example, in one of my tests I had 14% Proteobacteria, which often is elevated in people with gastric disorders like Crohn’s disease. That doesn’t mean that Proteobacteria is _bad_; in fact, it could mean the opposite, that we need more of these to fight back invaders. In my case the unusually high amount came right after a [trip to Central America](#experimentJungleTravel), during which my body was exposed to many new and possibly pathogenic microbes. My Proteobacteria fell soon after I returned to the U.S. and my normal diet.

Proteobacteria:  come back quickly after antibiotics (“a weed”) [QuantifiedBody](https://thequantifiedbody.net/human-microbiome-health-dr-rob-knight/)



Does Proteobacteria play a role?

TS: Only 52 of the 4200 compounds found in normal mice blood were identified in germ-free mice. The implication is that most of the chemicals in our blood are synthesized by microbes.

Using examples from my own sampling, can I say something about what to look for?

### Firmicutes

In most western guts, *Firmicutes* generally seems to dominate but, I find too much day-to-day variability to rely on a single test. Here's how my levels of the two most common phyla change on a daily basis:


```{r plotPhyla}

Firmicutes <- subset_taxa(subset_samples(sprague.phylum.norm, Site == "gut"), Phylum %in% c("Firmicutes","Bacteroidetes"))
fb <- mhg_abundance(subset_samples(Firmicutes, Date > "2017-03-15" & Date < "2017-03-27"))/10000 
fb <- fb["Firmicutes",]/fb["Bacteroidetes",] # 
fb <- fb %>% gather(Date,Ratio)
fb$Date <- as.Date(fb$Date)

ggplot(data=fb %>% gather(Date,Ratio),aes(x=Date,y=Ratio, group=1)) + geom_point() + geom_line() +
  scale_x_date(date_labels = "%m-%d (%A)", date_breaks = "1 day", date_minor_breaks = "1 day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Ten Consecutive Days of Firmicutes/Bacteroidetes")

# plot_bar(subset_samples(Firmicutes, Date > "2017-03-15" & Date < "2017-03-27"), fill = "Phylum", x = "Date") +
  # ggtitle("Firmicutes Abundance (%)") + scale_y_continuous(labels=function(x)x/10000) + ylab("Abundance (%)")

```


If you sampled on a Monday, you could well get a different answer than if you sampled on a Tuesday.

In fact, although there _does_ seem to be an average where my _Firmicutes_ is most abundant, there are many days where the amounts drop well below half.

```{r plotHistogramFirmicutes}

qplot(as.vector(otu_table(Firmicutes)[2,]/10000), geom="histogram", main="Histogram of Firmicutes Abundance", xlab = "Percentage of All My Samples", binwidth=5, fill=I("blue"), col = I("black"))

```


### Bacteroides

Although many of the species in this genus are pathogenic, you're much more likely to encounter one of the "mutualist" versions that are co-dependent on humans, providng us with many important features we have not had to evolve, while our gut gives back nutrients and more that _they_ need to survive. Read much more in [@wexler_bacteroides_2007].

### Bifidobacterium

The blue dots indicate days on which I ate some beans.

```{r exploreMicrobesBifido}

food<-rikfood
beef <- "beef|steak|ribeye|burger"
#beef.day <- unique(food[str_detect(food$Name, beef),"Date"])
beans <- "beans"
#beans.day <- unique(food[str_detect(food$Name, beans),"Date"])
# plot_bar(prune_taxa("Bifidobacterium",subset_samples(gut.norm, Site =="gut" & Date>"2016-07-01")), x = "Date") + scale_y_continuous(labels=function(x)x/10000) +    ggtitle("My Bifidobacterium") + ylab("Abundance (%)") +
#   annotate("pointrange",x = as.Date(beef.day), y = 0, ymin = -0.003, ymax = 0, color = "blue", size = 0.1) + 

beans.day <- mhg_food_days(food,beans)

keytaxa.top <- subset_taxa(gut.norm.nodupes, Genus %in% c("Bifidobacterium") )
 
plot_bar(subset_samples(keytaxa.top,Date > "2017-06-01" & Date < "2017-08-01" ), fill = "Genus", x = "Date") +
  ggtitle("Bifidobacterium Abundance (%)") + scale_y_continuous(labels=function(x)x/10000) + ylab("Abundance (%)") + 
  annotate("pointrange",x = beans.day[beans.day>as.Date("2017-06-01")], y = 0, ymin = -0.003, ymax = 0, color = "blue", size = 0.1) + 
  scale_x_date(date_labels = "%b20%y",
                date_breaks = "1 month", date_minor_breaks = "2 weeks")

```


### Lactobacillus

A 2017 careful study of rats revealed higher levels of _Lactobacillus_ on a diet of chicken (versus beef or soy)^[@zhu_beef_2017].  The blue dots represent days on which I ate poultry (chicken or turkey).  Green dots are days I ate beef.

```{r exploreMicrobesLacto}

food<-rikfood
chicken <- "chicken|turkey"
beef.day <- mhg_food_days(food,"beef|steak|ribeye")
beans <- "beans"
#beans.day <- unique(food[str_detect(food$Name, beans),"Date"])
# plot_bar(prune_taxa("Bifidobacterium",subset_samples(gut.norm, Site =="gut" & Date>"2016-07-01")), x = "Date") + scale_y_continuous(labels=function(x)x/10000) +    ggtitle("My Bifidobacterium") + ylab("Abundance (%)") +
#   annotate("pointrange",x = as.Date(beef.day), y = 0, ymin = -0.003, ymax = 0, color = "blue", size = 0.1) + 

beans.day <- mhg_food_days(food,beans)
chicken.day <- mhg_food_days(food,beans)
yogurt.day <- mhg_food_days(food,"yogurt")
kefir.day <- mhg_food_days(food,"kefir")

keytaxa.top <- subset_taxa(gut.norm, Genus %in% c("Lactobacillus") )
  
food.day <- chicken.day
food2.day <- beef.day

df <- rbind(data.frame(Date = as.Date(food.day),Food = "Chicken", Y=0),
            data.frame(Date = as.Date(mhg_food_days(food,beef)), Food = "Beef", Y=1)) %>% filter(Date > "2017-06-01")


plot_bar(subset_samples(keytaxa.top,Date > "2017-06-01" ), fill = "Genus", x = "Date") +
  ggtitle("Lactobacillus Abundance (%)") + scale_y_continuous(labels=function(x)x/10000) + ylab("Abundance (%)") + 
#  annotate("pointrange",x = food.day[food.day>as.Date("2017-04-01")], y = 0, ymin = -0.003, ymax = 0, color = "blue", size = 0.1) + 
 # annotate("pointrange",x = food2.day[food2.day>as.Date("2017-04-01")], y = 0, ymin = -0.003, ymax = 0, color = "green", size = 0.1) + 
  scale_x_date(date_labels = "%b20%y",
                date_breaks = "1 month", date_minor_breaks = "2 weeks") +
  xlim(as.Date("2017-06-01"),as.Date("2017-08-01")) +
  geom_point(data=df, aes(x=Date,y=0.03, color = Food), size=2, alpha = 0.8, inherit.aes = FALSE) +
  #geom_segment(data = df,aes(x=Date, xend=Date+1, y=0, yend=300, color=Food), size=10, inherit.aes = FALSE) +
 #annotate("pointrange",x = food2.day[food2.day>as.Date("2017-06-01")], y = 0, ymin = -0.003, ymax = 0, color = "green", size = 01) 
  labs(color = "Days when I ate")

#ggplot()+geom_segment(data = df,aes(x=Date, xend=Date, y=0-Y/10, yend=(0-Y/10)+.1, color=Food), size=1, inherit.aes = FALSE)
```

### Fat and my levels of Lactobacillus

Here's how much fat I eat daily.  Does this correlate with my levels of _Lactobacillus_?

```{r exploreMicrobesFat}

# unique(food[str_detect(food[["Fat"]], food),"Date"])
# 
# unique(food[food$Fat,"Date"])
fatPerDay <- food %>% group_by(Date) %>% summarise(sum(Fat)) # df with Date and total Fat for that day

# plot_bar(subset_samples(keytaxa.top,Date > "2017-01-01" ), fill = "Genus", x = "Date") +
#   ggtitle("Lactobacillus Abundance (%)") + scale_y_continuous(labels=function(x)x/10000) + ylab("Abundance (%)") # + geom_bar(data=fatPerDay, aes(x=Date,y=`sum(Fat)`),stat="identity")


# how much fat I ate
d = data.frame(Date=sample_data(keytaxa.top)$Date,Abundance=as.numeric(otu_table(keytaxa.top)["Lactobacillus",]))
ggplot() + geom_bar(data=fatPerDay, aes(x=Date,y=`sum(Fat)`),stat="identity", color = "red") + 
  scale_y_continuous(sec.axis = dup_axis())

#mhg_taxa(sprague.genus.norm,"Lactobacillus")
#grep("Lacto",taxa_names(sprague.species.norm), value=TRUE)

 #  geom_bar(color="blue",data=d, aes(x=Date,y=Abundance), stat = "identity")
 #                                                                                                        
 # range(fatPerDay$`sum(Fat)`)                                                                                                       
 #                                                                                                        
# 
# 
# d=data.frame(Date=sample_data(keytaxa.top)$Date,Abundance=as.numeric(otu_table(keytaxa.top)["Lactobacillus",]))
# 
# as.numeric(otu_table(keytaxa.top)["Lactobacillus",])

```




### Other microbes and digestion

Bacteroides plebeius : Japanese version can digest seaweed, thanks to genes stolen from Zobellia galactanivorans, that lives on the seaweed. @hehemann_transfer_2010

Notice that I seem to have quite a bit of this taxa:

```{r}
mhg_taxa_plot(subset_samples(sprague.species.norm, Site == "gut" & Date > "2017-01-01"), "Bacteroides plebeius")

taxa_names(subset_taxa(sprague.species.norm,grepl("Bacteroides",Species)))


```

Why?  and is it a coincidence that very few others I've tested have this taxa?  Is it also strange that I lived in Japan for about 10 years?  Could all that living in Japan have caused this microbe to take hold in me but not in other North Americans.



Renouf J Nutr 2011 says it's Bacteroides uniformis that breaks down the soy (and why Asians find soy healthier than americans)




